### 2022-04-11 pre/pre.tex
more

Let us give some intuition behind how these rules are applied in practice. The
\infname{ER} rule is used to replace equals by equals within a clause.
Consider the formula $\iforall x.\,  x \eq \cst{2} \iimplies \cst{even}(x)$. In
clausal form it is $X \noteq \cst{2} \llor \cst{even}(X)$. Applying the
\infname{ER} rule effectively applies the precondition $X \eq \cst{2}$ to the
conclusion $\cst{even}(X)$ to obtain $\cst{even}(\cst{2})$. The \infname{EF}
rule can be used to find instances of the clause that can be further simplified.
Consider the clause $\cst{f}(X) \eq \cst{a} \llor \cst{f}(Y) \eq \cst{a} \llor
\cst{p}(X, Y)$. Applying \infname{EF} to its first two literals yields $\cst{a}
\noteq \cst{a} \llor \cst{f}(X) \eq \cst{a} \llor \cst{p}(X, X)$ (with $\sigma = \{ Y \mapsto X \}$). This clause
then further simplifies to $\cst{f}(X) \eq \cst{a} \llor \cst{p}(X, X)$ using \infname{ER}.

>>>>

Let us give some intuition behind how these rules are applied in practice. The
\infname{ER} rule is used to establish reflexivity of equality---the property
that each term is equal to itself. Consider the formula $\iforall x.\,  x \eq
\cst{2} \iimplies \cst{even}(x)$. In clausal form it is $X \noteq \cst{2} \llor
\cst{even}(X)$. Applying the \infname{ER} rule effectively applies the
precondition $X \eq \cst{2}$ to the conclusion $\cst{even}(X)$ to obtain
$\cst{even}(\cst{2})$.

Similarly to the factoring rule in the standard resolution calculus, the
\infname{EF} rule can be used to find instances of the clause that duplicate
some literals.
Additionally, it takes care of some rarely occurring edge cases that are a consequence of restrictive
side-conditions of other rules of the calculus. Consider the clause $\cst{f}(X)
\eq \cst{a} \llor \cst{f}(Y) \eq \cst{a} \llor \cst{p}(X, Y)$. Applying
\infname{EF} to its first two literals yields $\cst{a} \noteq \cst{a} \llor
\cst{f}(X) \eq \cst{a} \llor \cst{p}(X, X)$ (with $\sigma = \{ Y \mapsto X \}$).
This clause then further simplifies to $\cst{f}(X) \eq \cst{a} \llor \cst{p}(X,
X)$ using \infname{ER}.
